QPLIB: A Library of Quadratic Programming Instances

### 2016-07-28 QPLIB-1.0.tex
implemented Nick's changes, a few additions (QP = MI-QCQP ...)

Quadratic Programming (QP) problems, where both the objective function and the constraints contain (at most) expressions in the variables of degree $2$, contain a surprisingly diverse set of rather different problems. According to the fine details of the formulation, solving a QP may require employing either fundamentally combinatorial techniques, or ideas rooted in nonlinear optimization principles, or a mix of the two. In this sense, QP is likely one of the classes of problems where the collaboration between the communities interested in combinatorial and nonlinear optimization is more necessary, and potentially fruitful.

However, this diversity also implies that QP means very different things to different researchers. It is perhaps therefore not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011}, so far there has never been a single library containing all different kinds of instances of QP. Several libraries devoted to special cases of QP are indeed available; however, each of them is either focussed on one application (a specific problem that can be modeled as QP), or on QPs with specific structural properties that make them suitable to be solved with some given class of algorithmic approaches. To the best of our knowledge,  collecting a set of instances of QP that is at the same time not overwhelmingly numerous and significant for the many different interested communities has not been attempted, yet. This work constitutes a first step in this direction.

In this paper we report the steps that have been done to collect a (hopefully) significant library of QP instances, filtering the large set of available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of all possible QP types. A particularly thorny issue in this process is how to select instances that are ``interesting''. Our solution is to take this to mean ``challenging for a significant set of solution methods''. Our filtering process has then been in part based on the idea that if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library. This also takes into account the fact that if very few (maybe one) solver can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance can still be deemed ``interesting''. Putting this approach in practice requires a nontrivial number of technical steps and decisions that are detailed in the paper. We hope that our work can provide useful guidelines for other interested researchers.

@@

Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research in the last decade. At least some of the rationale for this interest is likely due to the fact that they are the ``least nonlinear nonlinear problems''. Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP). This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others ({\bf citations?}). Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved. Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed ({\bf citations, like copositive/standard QP?}).

>>>>

Quadratic Programming (QP) problems, where both the objective function and the constraints contain (at most) expressions in the variables of degree $2$, include a surprisingly diverse set of rather different instances. According to the fine details of the formulation, solving a QP may require employing either fundamentally combinatorial techniques, or ideas rooted in nonlinear optimization principles, or a mix of the two. In this sense, QP is likely one of the classes of problems where the collaboration between the communities interested in combinatorial and nonlinear optimization is more necessary, and potentially fruitful.

However, this diversity also implies that QP means very different things to different researchers. This is illustrated by the simple fact that the class of problems that we simply refer to here as ``QP'' is called in different ways, among which Mixed-Integer Quadratically Constrained Quadratic Problems (MI-QCQP). It is perhaps therefore not surprising that, unlike for ``simpler'' problems classes \cite{Koch2011}, so far there has never been a single library containing all different kinds of instances of QP. Several libraries devoted to special cases of QP are indeed available; however, each of them is either focussed on one application (a specific problem that can be modeled as QP), or on QPs with specific structural properties that make them suitable to be solved with some given class of algorithmic approaches. To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly numerous and significant for the many different interested communities has not been attempted, yet. This work constitutes a first step in this direction.

In this paper we report the steps that have been done to collect a (hopefully) significant library of QP instances, filtering the large set of available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of all possible QP types. A particularly thorny issue in this process is how to select instances that are ``interesting''. Our choica has been to take this to mean ``challenging for a significant set of solution methods''. Our filtering process has then been in part based on the idea that if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library. Yet, if very few (maybe one) solvers can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance can still be deemed ``interesting''. Putting this approach in practice requires a nontrivial number of technical steps and decisions that are detailed in the paper. We hope that our work can provide useful guidelines for other interested researchers.

@@

Optimization problems with quadratic constraints and/or objective function (QP) have been the subject of a considerable amount of research over the last 60 years. At least some of the rationale for this interest is likely due to the fact that they are the ``least-nonlinear nonlinear problems''. Hence, in particular for the convex case, tools and techniques that have been honed during decades of research for Linear Programming (LP), typically with integrality constraints (MILP), can often be extended to the quadratic case with at least less effort than what would be required for tackling general Non Linear Programming (NLP) problems, without or with integrality constraints (MINLP). This has indeed happened over and over again with different algorithmic techniques, such as Interior Point methods, active-set methods (of which the simplex method is a prototypical example), enumeration methods, cut-generation techniques, reformulation techniques, and many others ({\bf citations?}). Similarly, nonconvex continuous QP are perhaps the ``simplest'' class of problems that require features like spatial enumeration techniques to be solved. Hence, they are both the natural basis for the development of general techniques for nonconvex NLP, and a very specific class so that specialized approaches can be developed ({\bf citations, like copositive/standard QP?}).
